---
title: "Hypothesis #2 - Higher spring water flows and possiblity of flooding exists when: spring temperatures are low"
output:
    html_notebook: default
    pdf_document: default
---

The purpose of this code is to determine if spring temperatures are related to flooding and/or to high water levels.

NOTE: Need to run the script 'Timing of Spring Melt' (SpMT function) and 'Streamflow' (SFMax function) first to get some of the variables needed to run this script.

Libraries to run when you open R
```{r}
library(CRHMr)
library(plyr)
library(tidyverse)
library(lubridate)
library(plotly)
library(reshape2)
library(tidyhydat)
library(weathercan)
library(reshape2)
library(scales)
library(zoo)
library(xts)

setwd("C:/Users/Goldmember/Documents/USask/MWSterm3project/R/FOR_PROJECT")
```

Variables (run first):
1. Enter the number of timesteps over which the mean (rolling average) is calculated, using the centered method (example: 27 [days] = average temperature using the 13 days before and 13 days after the 14th day - which in this script will be the day of year of spring melt initiation from Hypothesis 1)

2. Enter the site name from the Master.List to be evaluated
```{r}
#Number of timesteps over which the mean is calculated
RollAvg_4W = 27

#Name of the site
Site_Name = "Hay River"
```

Fucntion for calculating the four week temperature average, centered around the first day of spring melt, calculated previoulsy in Hypothesis 1 (SpMt.DOY). 

RUN LAST after all the variables and the function code have been run. SpMT = Spring Melt Temperature
```{r}
Spr4WAvg <- SprTmp(Master.List, RollAvg_4W, SpMT.DOY)
```

Code for running the four week average temperature function.
```{r}
SprTmp <- function (Master.List, RollAvg_4W, SpMT.DOY) {
   Master.List = Master.List[Master.List$Site == Site_Name ,]

  #Get the rolling average for the specified timestep in the variables, using the centered method
   Spr4WAvg<-Master.List %>%
      select('datetime','mt.1') %>%
      mutate(AvgT4W=rollmean(mt.1,RollAvg_4W, 
                             na.pad=TRUE, 
                             align="center") #mutate addes new column using rollmean output
             )   
   
   #Add a column for the day of year (DOY), and change all DOY formatting to numeric
   Spr4WAvg$DOY <- strftime(Spr4WAvg$datetime,format='%j')    
   Spr4WAvg$DOY <- as.numeric(Spr4WAvg$DOY)                    
   SpMT.DOY$DOY <- as.numeric(SpMT.DOY$DOY)

   #group by year
   Spr4WAvg <- Spr4WAvg %>% 
              group_by(year=year(datetime)
                       )
   
  #Joins the first day of spring for each year (SpMT.DOY) to the 27D avg dataset by year. So for example, in the year 1950, the same record from SpMT.DOY is pasted onto every record for Spr4WAvg that also occurs in 1950
   Spr4WAvg = data.frame(inner_join(Spr4WAvg,SpMT.DOY, by="year")) 

   #create a row that will be filled in later with the DOY of the spring melt timing from Hyp 1
   Spr4WAvg$check <- NA
   
   #Check each row to see if the DOY (DOY.x) is equal to the first DOY of spring melt (DOY.y, came from SpMT.DOY) for each year. When true, the DOY of spring melt is then pasted into the "check" column for later sorting.
   for (i in 1:nrow(Spr4WAvg)) {
      if (Spr4WAvg$DOY.x[i] == Spr4WAvg$DOY.y[i]){
      Spr4WAvg$check[i] <- Spr4WAvg$DOY.y[i]
      }
   }
   
   #Delete all the rows that contain NA in the check column, which singles out the DOY of spring melt
   Spr4WAvg <- Spr4WAvg[!is.na(Spr4WAvg$check),]
   
   #Select only the columns containing the date, DOY, and 27 day temperature average. The dataframe now contains the 27 day temperature average for the first day of spring every year. 
   Spr4WAvg<-Spr4WAvg %>% select('datetime.x','AvgT4W','DOY.x','year') 
   names(Spr4WAvg) <- c('datetime','AvgT4W','DOY','year')

   return(Spr4WAvg)
}
```

Graphically compare average spring temperature against peak streamflows
```{r}
#First, select a date range common to both datasets (flow and average spring temp) so that the row lengths are equal.
STemp <- Spr4WAvg[Spr4WAvg$datetime >= '1964-01-01' &
          Spr4WAvg$datetime <= '2014-12-31',]
SFSpr <- SFMax[SFMax$Date >= '1964-01-01' &
          SFMax$Date <= '2014-12-31',]

#Create linear model
fit_line <- lm(SFSpr$Value ~ STemp$AvgT4W)
summary(fit_line)
#P-value (signif is significant digits - the last digit)
PV <- signif(summary(fit_line)$coefficients[2,4],3)
#R-squared value
R2 <- signif(summary(fit_line)$r.squared,3)         

#plot
plot(STemp$AvgT4W,SFSpr$Value,
     xlab="Four week average temp centered around day of year of spring melt", ylab="Day of year of max streamflow")
abline(fit_line)
title(main="Average Spring Temperatures Plotted Against the Corresponding
   Maximum Spring Flows", 
   col.main="black", 
   font.main=4)
mtext(paste ('P-Value:',
             PV, sep = " ", 
             collapse = NULL),
   3, line=-2)
mtext(paste ('R-squared Value:',
             R2, 
             sep = " ", 
             collapse = NULL),
   3, line=-3)

```


Create linear regression model, including r-squared and p-value - WATER LEVEL SINCE 2002
```{r}
#For Level from 2002 on
SpTemp <- SpMT.DOY[SpMT.DOY$datetime>='1964-01-01' &
          SpMT.DOY$datetime<='2014-12-31',]
SFM <- SFMax[SFMax$Date>='1964-01-01' &
          SFMax$Date<='2014-12-31',]

STemp$DOY <- as.numeric(STemp$DOY)
SFM$DOY <- as.numeric(SFM$DOY)

#regression curve is called "fit_line" here using the linear model function
fit_line <- lm(SFM$DOY ~ STemp$DOY)
summary(fit_line)
PV <- signif(summary(fit_line)$coefficients[2,4],3) #P-value (signif is significant digits)
R2 <- signif(summary(fit_line)$r.squared,3)         #R-squared value


#plotting regression curve
plot(STemp$DOY, SFM$DOY,
     xlab="Day of year spring melt starts", ylab="Day of year of max streamflow")
fit_line <- lm(SFM$DOY ~ STemp$DOY)
abline(fit_line)
title(main="Spring Thaw Timing Plotted Against the Timing of 
   Maximum Spring Water Flows", col.main="black", font.main=4)
mtext(paste ('p-value:',PV, sep = " ", collapse = NULL),
   3, line=-2)
mtext(paste ('r-squared value:',R2, sep = " ", collapse = NULL),
   3, line=-3)
```


Part B
Are flood events assoicated with low spring temperatures?  Are spring temperatures during flood events above or below long-term average?

The flood years for the community of interest need to be entered in the 'FloodEvents' list.
```{r}
FloodEvents <- list(1952, 1963, 1965, 1972, 1974, 1978, 1979, 1981, 1982, 1985, 1986, 1989, 1992, 1994, 1997, 2001, 2003, 2005, 2006, 2007, 2008, 2009, 2010)
FloodEvents
#matches 
FE <- Spr4WAvg[Spr4WAvg$year %in% FloodEvents,]

NFE <- Spr4WAvg[!Spr4WAvg$year %in% FloodEvents,]

#mean four week spring temp for flood years
MFE <- mean(FE$AvgT4W, na.rm=TRUE)

#mean four week spring temp for all the years - excluding flood years
MNFE <- mean(NFE$AvgT4W , na.rm=TRUE)

MFE
MNFE
```


When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
