---
title: "Hypothesis #1 - Higher spring flows and possibility of flooding exists when: Spring Melt occurs late"
output:
    html_notebook: default
    pdf_document: default
---

The purpose of this code is to determine if the timing of spring melt is related to ice jam flooding.

NOTE: Need to run the script 'Timing of Spring Melt' (SpMT function) and 'Streamflow' (SFMax function) first to get some of the variables needed to run this script.

Libraries to run when you open R
```{r}
library(CRHMr)
library(plyr)
library(tidyverse)
library(lubridate)
library(plotly)
library(reshape2)
library(tidyhydat)
library(weathercan)
library(reshape2)
library(scales)
library(zoo)
library(xts)

setwd("C:/Users/Goldmember/Documents/USask/MWSterm3project/R/FOR_PROJECT")
```

First, select a date range common to both datasets - Spring Melt temp (SpTemp - derived from "Timing of Spring Melt" script, using the function name SpMT.DOY (for day of year of start of spring melt)) and Max Streamflow (SFMax - from the script "Streamflow".) Row lengths should be equal. Then, change the Day of Year (DOY) to numeric for calculations.
```{r}

SpTemp <- SpMT.DOY[SpMT.DOY$datetime>='1964-01-01' &
          SpMT.DOY$datetime<='2014-12-31',]
SFM <- SFMax[SFMax$Date>='1964-01-01' &
          SFMax$Date<='2014-12-31',]

STemp$DOY <- as.numeric(STemp$DOY)
SFM$DOY <- as.numeric(SFM$DOY)
```

Create linear regression model, including r-squared and p-value - STREAMFLOW
```{r}
#regression curve is called "fit_line" here using the linear model function
fit_line <- lm(SFM$DOY ~ STemp$DOY)
summary(fit_line)
PV <- signif(summary(fit_line)$coefficients[2,4],3) #P-value (signif is significant digits)
R2 <- signif(summary(fit_line)$r.squared,3)         #R-squared value


#plotting regression curve
plot(STemp$DOY, SFM$DOY,
     xlab="Day of Year Spring Melt Starts", ylab="Day of Year of Max Streamflow")
fit_line <- lm(SFM$DOY ~ STemp$DOY)
abline(fit_line)
title(main="Spring Thaw Timing Plotted Against the Timing of 
   Maximum Spring Water Flows", col.main="black", font.main=4)
mtext(paste ('p-value:',PV, sep = " ", collapse = NULL),
   3, line=-2)
mtext(paste ('r-squared value:',R2, sep = " ", collapse = NULL),
   3, line=-3)


```

Create linear regression model, including r-squared and p-value - WATER LEVEL SINCE 2002
```{r}
#For Level from 2002 on
SpTemp <- SpMT.DOY[SpMT.DOY$datetime >= '1964-01-01' &
          SpMT.DOY$datetime <= '2014-12-31',]
SFM <- SFMax[SFMax$Date >= '1964-01-01' &
          SFMax$Date <= '2014-12-31',]

STemp$DOY <- as.numeric(STemp$DOY)
SFM$DOY <- as.numeric(SFM$DOY)

#regression curve is called "fit_line" here using the linear model function
fit_line <- lm(SFM$DOY ~ STemp$DOY)
summary(fit_line)
PV <- signif(summary(fit_line)$coefficients[2,4],3) #P-value (signif is significant digits)
R2 <- signif(summary(fit_line)$r.squared,3)         #R-squared value


#plotting regression curve
plot(STemp$DOY, SFM$DOY,
     xlab="Day of year spring melt starts", ylab="Day of year of max streamflow")
fit_line <- lm(SFM$DOY ~ STemp$DOY)
abline(fit_line)
title(main="Spring Thaw Timing Plotted Against the Timing of 
   Maximum Spring Water Flows", col.main="black", font.main=4)
mtext(paste ('p-value:',PV, sep = " ", collapse = NULL),
   3, line=-2)
mtext(paste ('r-squared value:',R2, sep = " ", collapse = NULL),
   3, line=-3)
```


Part B
Are flood events associated with later spring melts? Is flood-related spring melt timing above or below the long term average?  This next part takes the mean of the flow column for non-flood years and the mean of the flow column for flood years and compares them.  

The flood years for the community of interest need to be entered in the 'FloodEvents' list.
```{r}
#Streamflows
FloodEvents <- list(1952, 1963, 1965, 1972, 1974, 1978, 1979, 1981, 1982, 1985, 1986, 1989, 1992, 1994, 1997, 2001, 2003, 2005, 2006, 2007, 2008, 2009, 2010)
FloodEvents
#matches 
FE <- SFM[SFM$year %in% FloodEvents,]

NFE <- SFM[!SFM$year %in% FloodEvents,]

#mean of flood year water level
MFE <- mean(FE$Value, na.rm=TRUE)

#mean of water level for all the years - excluding flood years
MNFE <- mean(NFE$Value , na.rm=TRUE)
```

```{r}
#Water Level
FloodEvents <- list(1952, 1963, 1965, 1972, 1974, 1978, 1979, 1981, 1982, 1985, 1986, 1989, 1992, 1994, 1997, 2001, 2003, 2005, 2006, 2007, 2008, 2009, 2010)
FloodEvents
#matches 
FE2 <- SFM[SFM$year %in% FloodEvents,]

NFE2 <- SFM[!SFM$year %in% FloodEvents,]

#mean of flood year water level
MFE2 <- mean(FE$Value, na.rm=TRUE)

#mean of water level for all the years - excluding flood years
MNFE2 <- mean(NFE$Value , na.rm=TRUE)
```


