---
title: "Hypothesis 5 - Higher spring water levels and possibility of flooding exists when: Rain-on-snow conditions occur"
output:
    html_notebook: default
    pdf_document: default
---

The purpose of this script is to determine if the occurance of rain-on-snow events are related to flooding and/or to high water levels.

NOTE: Need to run the scripts 'SpringMeltTiming' (SMT function) and 'Streamflow' (SFMax function) first

Libraries to run when you open R
```{r}
library(CRHMr)
library(plyr)
library(tidyverse)
library(lubridate)
library(plotly)
library(reshape2)
library(tidyhydat)
library(weathercan)
library(reshape2)
library(scales)
library(zoo)
library(xts)

setwd("C:/Users/Goldmember/Documents/USask/MWSterm3project/R/FOR_PROJECT")
```

Rain on snow function
```{r}
ROS <- Rain_Snow(Site_Name, Master.List)
```

Variables needed for function:
1. Call in the Master List if not already in the global environment
2. Fill in the site name for which community you are wanting to look at from the Master List.
3. Choose the date range for which you'd like to calculate the amount of rain on snow (DOYS = day of year to start with; DOYE = day of year to end with)
```{r}
#1
Master.List <- read.csv("C:/Users/Goldmember/Documents/USask/MWSterm3project/R/FOR_PROJECT", header = TRUE)

#2
Site_Name = "Hay River"

#3
DOYS = '074'   #March 15
DOYE = '181'  #June 30

```

Function for calculating 'Rain on Snow" amounts
```{r}

#Code for Spring Melt Function - selects the first occurance of 5 consecutive days with min temps > 0 degrees C in the spring months
  Rain_Snow <- function (Site_Name, Master.List) {
    Master.List = Master.List[Master.List$Site == Site_Name ,]
    Master.List$datetime <- as.POSIXct(Master.List$datetime)
   
#ROSData = Rain On Snow Data
   ROSData <- data.frame(Master.List$datetime, 
                         signif(Master.List$sg.1, 3), 
                        signif(Master.List$tr.1, 3))
   
#Select all records within a given DOY range (with DOYS and DOYE specified in variables)   
   names(ROSData) <- c('datetime','sg.1','tr.1')    
   ROSData$DOY <- strftime(ROSData$datetime, format='%j')
   ROSData <- ROSData[ROSData$DOY >= DOYS &
              ROSData$DOY <= DOYE,]

#Select only records with snowfall on the ground
   ROSData <- ROSData[ROSData$sg.1 > 0,]
   
#Sum the amount of rainfall that occured when snow was present
   ROSSum <- ROSData %>% 
      group_by(year=year(datetime)) %>% 
      summarize(SWE = sum(tr.1, na.rm = TRUE)) 
 
#USE this if NA rows exist
#   ROSSum <- ROSSum[!is.na(ROSSum$year),]
   
   return(data.frame(ROSSum))
}

```

Linear model for comparing amount of rain on snow against peak springtime streamflow
```{r}
#Select a date range common to both datasets (so that the row lengths are equal)
R_O_S <- ROS[ROS$year >= '1964' &
          ROS$year <= '2014',]

SFSpr <- SFMax[SFMax$Date >= '1964-01-01' &
          SFMax$Date <= '2014-12-31',]

#Create linear model
fit_line <- lm(SFSpr$Value ~ R_O_S$SWE)
summary(fit_line)

#P-value (signif is significant digits - the last digit)
PV <- signif(summary(fit_line)$coefficients[2,4],3) 
#R-squared value
R2 <- signif(summary(fit_line)$r.squared,3)         

#plot
plot(R_O_S$SWE,SFSpr$Value,
     xlab="Rain on snow events", ylab="Day of year of max streamflow")
abline(fit_line)
title(main="Average Spring Temperatures Plotted Against the Corresponding
   Maximum Spring Water Levels", col.main="black", font.main=4)
mtext(paste ('P-Value:',PV, sep = " ", collapse = NULL),
   3, line=-2)
mtext(paste ('R-squared Value:',R2, sep = " ", collapse = NULL),
   3, line=-3)

```


Linear model for comparing amount of rain on snow against peak WATER LEVEL
```{r}
#Select a date range common to both datasets (so that the row lengths are equal)
R_O_S <- ROS[ROS$year >= '1964' &
          ROS$year <= '2014',]

SFSpr <- SFMax[SFMax$Date >= '1964-01-01' &
          SFMax$Date <= '2014-12-31',]

#Create linear model
fit_line <- lm(SFSpr$Value ~ R_O_S$SWE)
summary(fit_line)

#P-value (signif is significant digits - the last digit)
PV <- signif(summary(fit_line)$coefficients[2,4],3) 
#R-squared value
R2 <- signif(summary(fit_line)$r.squared,3)         

#plot
plot(R_O_S$SWE,SFSpr$Value,
     xlab="Rain on snow events", ylab="Day of year of max streamflow")
abline(fit_line)
title(main="Average Spring Temperatures Plotted Against the Corresponding
   Maximum Spring Water Levels", col.main="black", font.main=4)
mtext(paste ('P-Value:',PV, sep = " ", collapse = NULL),
   3, line=-2)
mtext(paste ('R-squared Value:',R2, sep = " ", collapse = NULL),
   3, line=-3)

```


Part B
Are flood events associated with rain-on-snow events? Are rain-on-snow events that occur prior to ice jam floods above or below long-term average?

The flood years for the community of interest need to be entered in the 'FloodEvents' list.
```{r}
FloodEvents <- list(1952, 1963, 1965, 1972, 1974, 1978, 1979, 1981, 1982, 1985, 1986, 1989, 1992, 1994, 1997, 2001, 2003, 2005, 2006, 2007, 2008, 2009, 2010)
FloodEvents
#matches 
FE <- R_O_S[R_O_S$year %in% FloodEvents,]

NFE <- R_O_S[!R_O_S$year %in% FloodEvents,]

#mean two week peak October streamflow for flood years
MFE <- mean(FE$SWE, na.rm=TRUE)

#mean two week peak October streamflow for all the years - excluding flood years
MNFE <- mean(NFE$SWE , na.rm=TRUE)

MFE
MNFE

```

