---
title: "'Hypothesis #3 - Higher spring water levels and the possibility of flooding"
  exists when: Snowpack SWE at the end of winter is high'
output:
  pdf_document: default
  word_document: default
---

The purpose of this code is to determine if the snowpack SWE is related to flooding and/or high water levels that lead to ice jam flooding. A majority of this code was adapted from code previously developed by Dr. Ryan Connon, Hydrologist for the Government of the Northwest Territories.

NOTE: Need to run the script 'Timing of Spring Melt' (SpMT function) and 'Streamflow' (SFMax function) first to get some of the variables needed to run this script.

Libraries to run when you open R
```{r}
library(tidyhydat)
library(ggplot2)
library(Kendall)
library(dplyr)
library(lubridate)
library(plotly)
library(xts)
library(mblm)

setwd("C:/Users/Goldmember/Documents/USask/MWSterm3project/R/FOR_PROJECT")

Master.List <- read.csv("C:/Users/Goldmember/Documents/USask/MWSterm3project/R/FOR_PROJECT/Master.List.csv", header = TRUE)


```

Function to analyze snow water equivalent (SWE) at each site
RUN LAST
```{r}
SWE <- Winter.SWE(Master.List, 
                site_name = "Aklavik",
                days_of_record = 200,
                start_year = 1850,
                end_year = 2050,
                start_month = 10, #inclusive
                end_month = 4)

```

Code for runnning the function "SWE" (from Ryan Connon)
```{r}
# Start and end dates can be selected based on what you are looking for, also update site_name as needed.

Winter.SWE <- function(Master.List, 
                       site_name = "Aklavik",
                       days_of_record = 200,
                       start_year = 1850,
                       end_year = 2050,
                       start_month = 10, #inclusive
                       end_month = 4) #inclusive 
{
    if (start_month == 1) {
      start <- "01 Jan"
    } else if (start_month == 2) {
      start <- " 01 Feb"
    } else if (start_month == 3) {
      start <- "01 Mar"
    } else if (start_month == 4) {
      start <- "01 Apr"
    } else if (start_month == 5) {
      start <- "01 May"
    } else if (start_month == 6) {
      start <- "01 Jun"
    } else if (start_month == 7) {
      start <- "01 Jul"
    } else if (start_month == 8) {
      start <- "01 Aug"
    } else if (start_month == 9) {
      start <- "01 Sep"
    } else if (start_month == 10) {
      start <- "01 Oct"
    } else if (start_month == 11) {
      start <- "01 Nov"
    } else if (start_month == 12) {
      start <- "01 Dec"
    }
    
    if (end_month == 1) {
      end <- "31 Jan"
    } else if (end_month == 2) {
      end <- "28 Feb"
    } else if (end_month == 3) {
      end <- "31 Mar"
    } else if (end_month == 4) {
      end <- "30 Apr"
    } else if (end_month == 5) {
      end <- "31 May"
    } else if (end_month == 6) {
      end <- "30 Jun"
    } else if (end_month == 7) {
      end <- "31 Jul"
    } else if (end_month == 8) {
      end <- "31 Aug"
    } else if (end_month == 9) {
      end <- "30 Sep"
    } else if (end_month == 10) {
      end <- "31 Oct"
    } else if (end_month == 11) {
      end <- "30 Nov"
    } else if (end_month == 12) {
      end <- "31 Dec"
    }
    
  wtr_yr <- function(dates, 
                     start_month = 10) 
    {
    dates.posix = as.POSIXlt(dates)
    offset = ifelse(dates.posix$mon >= start_month - 1, 1, 0)
    adj.year = dates.posix$year + 1900 + offset
    adj.year
  }

#Called in the Master List
data = Master.List[Master.List$Site == site_name ,]

#Rename columns to match Ryans Code
names(data) <- c('site',
                  'date',
                  'mean_temp', 
                  'max_temp', 
                  'min_temp', 
                  'tsnow_ground', 
                  'total_rain_mm', 
                  'total_snow',
                  'total_precip_mm')

Summary <- data %>%
  mutate(water_year = wtr_yr(data$date)) %>%
  mutate(month = month(date)) %>%             #added a month column
  filter(water_year >= start_year) %>%
    filter(water_year <= end_year) %>%
    filter(month <= end_month | month >= start_month) %>%
    group_by(water_year) %>%
    summarize(SWE = sum(total_precip_mm, na.rm = TRUE) - sum(total_rain_mm, na.rm = TRUE),
              n1 = sum(!is.na(total_precip_mm)),
              n2 = sum(!is.na(total_rain_mm))) %>%
  mutate(n = ifelse(n1 >= n2, n1, n2)) %>%
  select(-n1, -n2) %>%
  subset(water_year >= start_year) %>%
  subset(water_year <= end_year) %>%
  subset(n >= days_of_record)

  Summary <- data.frame(Summary)
  
  Summary$SWE[Summary$SWE<0] <- NA

  return(Summary)
}

```


Linear Regression model for comparing maximum SWE against peak streamflow
```{r}
#Select a date range common to both datasets (so that the row lengths are equal)
SWEMax <- SWE[SWE$water_year >= '1940' &
          SWE$water_year <= '2015',]
SFSpr <- SFMaxAk[SFMaxAk$Date >= '1939-01-01' &
          SFMaxAk$Date <= '2016-12-31',]

SWEMax <- SWE[SWE$water_year >= 1974 &
                 SWE$water_year <= 1977 |
                 SWE$water_year >= 2009 &
                 SWE$water_year <= 2016 ,]

SFSpr <- SFMaxAk[SFMaxAk$Date >= 1974 &
                 SFMaxAk$Date <= 1977 |
                 SFMaxAk$Date >= 2009 &
                 SFMaxAk$Date <= 2016 ,]

#Create linear regression model
fit_line <- lm(SFSpr$Value ~ SWEMax$SWE)
summary(fit_line)

#P-value (signif is significant digits - the last digit)
PV <- signif(summary(fit_line)$coefficients[2,4],3)
#R-squared value
R2 <- signif(summary(fit_line)$r.squared,3)        

#plot
plot(SWEMax$SWE,SFSpr$Value,
     xlab="Maximum annual SWE", ylab="Maximum streamflow")
abline(fit_line)
title(main="Max Snow Water Equivalent Plotted Against the Corresponding
   Maximum Spring Streamflows for Aklavik", col.main="black", font.main=4)
mtext(paste ('P-Value:',PV, sep = " ", collapse = NULL),
   3, line=-2)
mtext(paste ('R-squared Value:',R2, sep = " ", collapse = NULL),
   3, line=-3)

```



Linear Regression model for comparing maximum SWE against peak WATER LEVEL
```{r}
#Select a date range common to both datasets (so that the row lengths are equal)
SWEMax <- SWE[SWE$water_year >= '2002' &
          SWE$water_year <= '2016',]
WLSpr <- WLMaxFS[WLMaxFS$Date >= '2002-01-01' &
          WLMaxFS$Date <= '2017-12-31',]

#Create linear regression model
fit_line <- lm(WLSpr$Value ~ SWEMax$SWE)
summary(fit_line)

#P-value (signif is significant digits - the last digit)
PV <- signif(summary(fit_line)$coefficients[2,4],3)
#R-squared value
R2 <- signif(summary(fit_line)$r.squared,3)        

#plot
plot(SWEMax$SWE,WLSpr$Value,
     xlab="Maximum annual SWE", ylab="Maximum Water Levels")
abline(fit_line)
title(main="Max Snow Water Equivalent Plotted Against the Corresponding
   Maximum Spring Water Levels for Fort Simpson", col.main="black", font.main=4)
mtext(paste ('P-Value:',PV, sep = " ", collapse = NULL),
   3, line=-2)
mtext(paste ('R-squared Value:',R2, sep = " ", collapse = NULL),
   3, line=-3)
```


Part B
Are flood events assoicated with high SWE amounts? Are SWE amounts during flood events above or below long-term average?

The flood years for the community of interest need to be entered in the 'FloodEvents' list.
```{r}
# FloodEvents <- list(1952, 1963, 1965, 1972, 1974, 1978, 1979, 1981, 1982, 1985, 1986, 1989, 1992, 1994, 1997, 2001, 2003, 2005, 2006, 2007, 2008, 2009, 2010)
FloodEvents <- list(1961, 1982, 1992, 2006, 2013)
FloodEvents
#matches 
FE <- SWEMax[SWEMax$water_year %in% FloodEvents,]

NFE <- SWEMax[!SWEMax$water_year %in% FloodEvents,]

#mean four week spring temp for flood years
MFE <- mean(SWEMax$SWE, na.rm=TRUE)

#mean four week spring temp for all the years - excluding flood years
MNFE <- mean(NFE$SWE , na.rm=TRUE)

MFE
MNFE


```

