---
title: "Hypothesis 4 - Higher spring water levels and possibility of flooding exists when: Water levels were high in the previous fall."  
output:
    html_notebook: default
    pdf_document: default
---

The purpose of this code is to determine if fall water levels are related to spirng flooding and/or to high spring water levels 

NOTE: Need to run the script 'Streamflow' (SFMax function) first to get some of the variables needed to run this script. Fall water levels (example, the average water level in a two week window in October) was determined in the Streamflow script. So this script just calls that dataframe.

Libraries to run when you open R
```{r}
library(CRHMr)
library(plyr)
library(tidyverse)
library(lubridate)
library(plotly)
library(reshape2)
library(tidyhydat)
library(weathercan)
library(reshape2)
library(scales)
library(zoo)
library(xts)

setwd("C:/Users/Goldmember/Documents/USask/MWSterm3project/R/FOR_PROJECT")
```

Linear regression model for comparing peak fall streamflow against peak spring streamflow.
First, select a date range common to both datasets (so that row lengths are equal).
```{r}
#Streamflow Fall
SFFall <- SFFall_MaxFS[SFFall_MaxFS$Date >= '1964-01-01' &
          SFFall_MaxFS$Date <= '2014-12-31',]
#Streamflow Spring
SFSpr <- SFMaxFS[SFMaxFS$Date >= '1964-01-01' &
          SFMaxFS$Date <= '2014-12-31',]

#Create linear model
fit_line <- lm(SFSpr$Value ~ SFFall$Value)
summary(fit_line)
PV <- signif(summary(fit_line)$coefficients[2,4],3) #P-value (signif is significant digits - the last digit)
R2 <- signif(summary(fit_line)$r.squared,3)         #R-squared value

#plot
plot(SFFall$Value,SFSpr$Value,
       xlab="Maximum fall streamflows", ylab="Maximum spring streamflow")
abline(fit_line)
title(main="Peak Fall Streamflow Plotted Against the Corresponding
   Peak Spring Streamflow for Fort Simpson", col.main="black", font.main=4)
mtext(paste ('P-Value:',PV, sep = " ", collapse = NULL),
   3, line=-2)
mtext(paste ('R-squared Value:',R2, sep = " ", collapse = NULL),
   3, line=-3)

```

Linear regression model for comparing peak fall streamflow against peak WATER LEVEL.
First, select a date range common to both datasets (so that row lengths are equal).
```{r}
#Water Level Fall
WLFall <- WLFall_MaxFS[WLFall_MaxFS$Date >= '1964-01-01' &
          WLFall_MaxFS$Date <= '2014-12-31',]
#Water Level Spring
WLSpr <- WLMaxFS[WLMaxFS$Date >= '1964-01-01' &
          WLMaxFS$Date <= '2014-12-31',]

#Create linear model
fit_line <- lm(WLSpr$Value ~ WLFall$Value)
summary(fit_line)
PV <- signif(summary(fit_line)$coefficients[2,4],3) #P-value (signif is significant digits - the last digit)
R2 <- signif(summary(fit_line)$r.squared,3)         #R-squared value

#plot
plot(WLFall$Value,WLSpr$Value,
       xlab="Maximum fall water levels", ylab="Maximum spring water levels")
abline(fit_line)
title(main="Peak Fall Streamflow Plotted Against the Corresponding
   Peak Spring Water Levels for Fort Simpson", col.main="black", font.main=4)
mtext(paste ('P-Value:',PV, sep = " ", collapse = NULL),
   3, line=-2)
mtext(paste ('R-squared Value:',R2, sep = " ", collapse = NULL),
   3, line=-3)



```





Part B
Are flood events assoicated with high fall water levels from the previous year? Are fall streamflows prior to ice jam floods above or below long-term average?

The flood years for the community of interest need to be entered in the 'FloodEvents' list.
```{r}
# FloodEvents <- list(1952, 1963, 1965, 1972, 1974, 1978, 1979, 1981, 1982, 1985, 1986, 1989, 1992, 1994, 1997, 2001, 2003, 2005, 2006, 2007, 2008, 2009, 2010)
FloodEvents <- list(1961, 1982, 1992, 2006, 2013)
FloodEvents
#matches 
FE <- SFFall_MaxAk[SFFall_MaxAk$year %in% FloodEvents,]

NFE <- SFFall_MaxAk[!SFFall_MaxAk$year %in% FloodEvents,]

#mean two week peak October streamflow for flood years
MFE <- mean(FE$AvgT2W, na.rm=TRUE)

#mean two week peak October streamflow for all the years - excluding flood years
MNFE <- mean(NFE$AvgT2W , na.rm=TRUE)

MFE
MNFE

```
Flood years vs. water level
```{r}
FloodEvents <- list(1952, 1963, 1965, 1972, 1974, 1978, 1979, 1981, 1982, 1985, 1986, 1989, 1992, 1994, 1997, 2001, 2003, 2005, 2006, 2007, 2008, 2009, 2010)
FloodEvents
#matches 
FE2 <- WLFall_Max[WLFall_Max$year %in% FloodEvents,]

NFE2 <- WLFall_Max[!WLFall_Max$year %in% FloodEvents,]

#mean two week peak October streamflow for flood years
MFE2 <- mean(FE2$AvgT2W, na.rm=TRUE)

#mean two week peak October streamflow for all the years - excluding flood years
MNFE2 <- mean(NFE2$AvgT2W , na.rm=TRUE)

MFE2
MNFE2

```