---
title: "Day of year of maximum annual streamflow; all streamflows; and max fall streamflows "
output:
    html_notebook: default
    pdf_document: default
---

This script will provide the day of year (DOY) of the maximum annual streamflow, all the streamflows for a given station during a given time period, and/or the max annual streamflow for a specified station.

```{r}
library(CRHMr)
library(plyr)
library(tidyverse)
library(lubridate)
library(plotly)
library(reshape2)
library(tidyhydat)
library(weathercan)
library(reshape2)
library(scales)
library(zoo)
library(xts)

setwd("C:/Users/Goldmember/Documents/USask/MWSterm3project/R/FOR_PROJECT")
```

Streamflow function.  Run LAST after variables and function code have been run.
```{r}
SFMax <- StreamflowMax (STID1, DateS, DateE, DOY)  #Max Streamflow each year
SF <- Streamflow (STID1, DateS, DateE, DOY)        #All Streamflow
SFFall_Max <- StreamflowFall (STID1, DateS, DateE, DOYS, DOYE) #Max Fall Streamflow
```

Variables (run first):
1. Enter the station ID (STID1) number
2. Enter the start (DateS) and end (DateE) dates for the flow period of interst
3. Cutoff day of year (DOY) from which to select spring streamflows 
4. Cutoff days of year (DOYS = start day, DOYE = end day) for Hypothesis 4 (fall water levels).
```{r}
# 1
STID1 = '07OB001'

# 2
DateS = "1964-01-01"
DateE = "2017-12-31"

# 3 - example: 166 Juliene Days = June 15th; June 16th in a non leap year
DOY = 166

# 4 - Cutoff DOY from which to select fall streamflow for number 8 (H4)
DOYS = 274 #Oct 1st
DOYE = 304 #Oct 31st
#Number of timesteps over which the mean is calculated
RollAvg_2W = 15
```

Function code for maximum annual streamflow
```{r}
StreamflowMax <- function(STID1, DateS, DateE, DOY) {

   WF <- hy_daily_flows(station_number = STID1,
                       start_date = DateS, 
                       end_date = DateE
                       )

   WF$Date <- as.POSIXct(WF$Date)

   #Select all streamflow records before a given DOY
   WF$DOY <- strftime(WF$Date, format='%j')
   WF <- WF[WF$DOY <= DOY,]


  #Selecting the max streamflow each year
   WFmax <- WF %>% 
            group_by(year=year(Date)) %>% 
            slice (which.max(Value))

return(WFmax)
}
```

Function code for all streamflow from a specified station for a specified date range
```{r}
Streamflow <- function(STID1, DateS, DateE, DOY) {

   WF <- hy_daily_flows(station_number = STID1,
                       start_date = DateS, 
                       end_date = DateE
                       )

   WF$Date <- as.POSIXct(WF$Date)

   #Create a DOY column
   WF$DOY <- strftime(WF$Date, format='%j')

return(WF)
}
```

Function code for maximum fall streamflow
```{r}
StreamflowFall <- function(STID1, DateS, DateE, DOYS, DOYE) {

   WF <- hy_daily_flows(station_number = STID1,
                       start_date = DateS, 
                       end_date = DateE
                       )

   WF$Date <- as.POSIXct(WF$Date)

   #Select all streamflow records within a given DOY range
   WF$DOY <- strftime(WF$Date, format='%j')
   WF <- WF[WF$DOY >= DOYS &
            WF$DOY <= DOYE,]

   Fall2WAvg <- WF %>%
                select('Date','Value','DOY') %>%
                mutate(AvgT2W = rollmean(Value, RollAvg_2W, na.pad=TRUE, align="center")) 

   #Delete NAs
   Fall2WAvg <- Fall2WAvg[!is.na(Fall2WAvg$AvgT2W),]

  #Selecting the max streamflow each year
   Fall2WMax <- Fall2WAvg %>% 
                group_by(year=year(Date)) %>% 
                slice (which.max(AvgT2W))

   Fall2WMax <- data.frame(Fall2WMax)
return(Fall2WMax)
}

```

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
