---
title: "Day of year of maximum annual water level; all water levels (available from the Water Survey); and max fall water levels"
output:
    html_notebook: default
    pdf_document: default
---

This script will provide the day of year (DOY) of the maximum annual water level, all the water levels for a given station during a given time period, and/or the max annual water level for a specified station. This script should be run before others.

```{r}
library(CRHMr)
library(plyr)
library(tidyverse)
library(lubridate)
library(plotly)
library(reshape2)
library(tidyhydat)
library(weathercan)
library(reshape2)
library(scales)
library(zoo)
library(xts)

setwd("C:/Users/Goldmember/Documents/USask/MWSterm3project/R/FOR_PROJECT")
```

Water Levels function.  Run LAST after variables and function code have been run.
```{r}
WLMax <- Water.Level.Max (STID1, DateS, DateE, DOY)  #Max water level each year
WL <- Water.Level (STID1, DateS, DateE, DOY)        #All water levels
WLFall_Max <- Water.Level.Fall (STID1, DateS, DateE, DOYS, DOYE) #Max Fall water levels
```

Variables (run first):
1. Enter the station ID (STID1) number
2. Enter the start (DateS) and end (DateE) dates for the water level period of interst
3. Cutoff day of year (DOY) from which to select spring water levels
4. Cutoff days of year (DOYS = start day, DOYE = end day) for Hypothesis 4 (fall water levels).
```{r}
# 1
STID1 = '07OB001'

# 2
DateS = "1964-01-01"
DateE = "2017-12-31"

# 3 - example: 166 Juliene Days = June 15th; June 16th in a non leap year
DOY = 166

# 4 - Cutoff DOY from which to select fall water level for Hypothesis 4)
DOYS = 274 #Oct 1st
DOYE = 304 #Oct 31st
#Number of timesteps over which the mean is calculated
RollAvg_2W = 15
```

Function code for maximum annual water levels
```{r}
Water.Level.Max <- function(STID1, DateS, DateE, DOY) {

   WL <- hy_daily_levels(station_number = STID1,
                       start_date = DateS, 
                       end_date = DateE
                       )

   WL$Date <- as.POSIXct(WL$Date)

   #Select all water level records before a given DOY
   WL$DOY <- strftime(WL$Date, format='%j')
   WL <- WL[WL$DOY <= DOY,]


  #Selecting the max streamflow each year
   WLmax <- WL %>% 
            group_by(year=year(Date)) %>% 
            slice (which.max(Value))

return(WLmax)
}
```

Function code for all water levels from a specified station for a specified date range
```{r}
Water.Level <- function(STID1, DateS, DateE, DOY) {

   WL <- hy_daily_levels(station_number = STID1,
                       start_date = DateS, 
                       end_date = DateE
                       )

   WL$Date <- as.POSIXct(WL$Date)

   #Create a DOY column
   WL$DOY <- strftime(WL$Date, format='%j')

return(WL)
}
```

Function code for maximum fall streamflow
```{r}
Water.Level.Fall <- function(STID1, DateS, DateE, DOYS, DOYE) {

   WL <- hy_daily_levels(station_number = STID1,
                       start_date = DateS, 
                       end_date = DateE
                       )

   WL$Date <- as.POSIXct(WL$Date)

   #Select all water level records within a given DOY range
   WL$DOY <- strftime(WL$Date, format='%j')
   WL <- WL[WL$DOY >= DOYS &
            WL$DOY <= DOYE,]

   Fall2WAvg <- WL %>%
                select('Date','Value','DOY') %>%
                mutate(AvgT2W = rollmean(Value, RollAvg_2W, na.pad=TRUE, align="center")) 

   #Delete NAs
   Fall2WAvg <- Fall2WAvg[!is.na(Fall2WAvg$AvgT2W),]

  #Selecting the max streamflow each year
   Fall2WMax <- Fall2WAvg %>% 
                group_by(year=year(Date)) %>% 
                slice (which.max(AvgT2W))

   Fall2WMax <- data.frame(Fall2WMax)
return(Fall2WMax)
}

```

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
